Cryptanalysis of SBC based post-quantum signature scheme

### 2025-05-27 rapport.tex
phase 1 presque terminée

% !TXS template
\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[a4paper]{geometry}
\usepackage{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage[backend=bibtex]{biblatex}
\usepackage{csquotes}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{hyperref}

@@

\appendix
\section{Signature protocol}

@@

	\begin{theorem}
		Suppose that no adversary can distinguish the PRF used by the protocol from real random bits
		$$
		| \mathcal{P}_{seed \in \{0,1\}^n}[b=0 | p \leftarrow PRF(seed)] - \mathcal{P}[b=1 | p \overset{{\scriptscriptstyle\$}}{\leftarrow} \{0,1\}^{2n}] | \leq \frac{1}{2}.
		$$
		Suppose also that the protocol uses a cryptographic hash function which is one-way and does not have collision attacks with better complexity than $\frac{1}{\lambda}$ using a $\mathcal{PPT}$, where $\lambda$ is the number security bits needed for the signature scheme.
		Finally, suppose that the SBC problem is hard, i.e. the complexity the best attack of the SBC problem has complexity $\frac{1}{\lambda}$ using a $\mathcal{PPT}$.
		Then, the SBC signature scheme is \textbf{EUF-CMA}.
	\end{theorem}
	\begin{proof}
		\textbf{Il suffit de modeler le schema avec les énoncés de \cite{feneu2023} et hop on peut simplement utiliser ses preuves qui sont assez gnérales pour tout système de signature MPCitH avec une oneway function et une PRF.}
	\end{proof}

>>>>

% !TXS template
\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[a4paper]{geometry}
\usepackage{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage[backend=bibtex]{biblatex}
\usepackage{csquotes}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{hyperref}
\usepackage{tcolorbox}
\usepackage{enumitem}
\usepackage{fancyvrb}
\usepackage{multicol}
\usepackage{stmaryrd}

@@

\appendix
\section{Signature protocol}
	Here we will show to signature protocol adapted from the identification protocol from \cite{HJ23} in the same format as in \cite{feneu2023} as a 5-round protocol to be able to apply his theorem on EUF-CMA security proof of such protocol.

		\begin{tcolorbox}[title=Signature Scheme -- Signing algorithm, colback=white, colframe=black]
			\textbf{Inputs:} A public key $x = (\vec{u}, \vec{v})$, a private key $w = (\vec{x}, \vec{y}) \in \text{NSBC}[\vec{u}, \vec{v}]$ and a message $m \in \{0,1\}^*$.

			\vspace{0.5em}
			\textbf{Phase 0: Initialization.}
			\begin{enumerate}[label=\arabic*.]
				\item Sample a random salt $\texttt{salt} \leftarrow \{0,1\}^{\lambda}$.
				\item Sample a root seed $\rho \leftarrow \{0,1\}^{\lambda}$.
			\end{enumerate}

			\vspace{0.5em}
			\textbf{Phase 1: Preparation of the MPC-in-the-Head inputs.} For each iteration $e \in [1:\tau]$:
			\begin{enumerate}[label=\arabic*.]
				\item Derive randomness $r^{[e]}_{w}, r^{[e]}_r, r^{[e]}_{\psi}$ from the root seed $\rho$: \\
				\quad $(r^{[e]}_w, r^{[e]}_r, r^{[e]}_{\psi}) \leftarrow \text{PRG}(\texttt{salt}, e, \rho)$.

				\item Share the witness $w$ into an $(N, N)$-threshold linear secret sharing $[w^{[e]}_i]$: \\
				\quad $[X_1^{[e]}], [X_2^{[e]}], [Y_1^{[e]}], [Y_2^{[e]}], [R_A^{[e]}], [R_B^{[e]}] \leftarrow \text{Share}(w; r^{[e]}_w)$ \#The shares are in $\mathbb{F}_{q^k}$. \\
				\quad $[\vec{R}_y^{[e]}], [\vec{R}_y^{[e]}] \leftarrow \text{Share}(w; r^{[e]}_w)$ \#The shares are in $(\mathbb{F}_{q})^n$.

				\item Compute $X_1, X_2, Y_1, Y_2$ \\
				\quad $X_1 \leftarrow \sum_{i=1}^{N} [X_1^{[e]}]^{\llbracket i \rrbracket}, X_2 \leftarrow \sum_{i=1}^{N} [X_2^{[e]}]^{\llbracket i \rrbracket}, Y_1 \leftarrow \sum_{i=1}^{N} [Y_1^{[e]}]^{\llbracket i \rrbracket}, Y_2 \leftarrow \sum_{i=1}^{N} [Y_2^{[e]}]^{\llbracket i \rrbracket}$

				\item Compute $A^{[e]}$ and $B^{[e]}$ and share them: \\
				\quad $A \leftarrow X_1Y_1 - X_2Y_2$. \\
				\quad $B \leftarrow X_1(\vec{v} \cdot \vec{y}) + Y_1(\vec{u} \cdot \vec{x}) - X_2(\vec{u} \cdot \vec{y}) - Y_2(\vec{v} \cdot \vec{x})$. \\
				\quad $\delta_A \leftarrow A - \sum_{i=1}^{N} [R_A^{[e]}]^{\llbracket i \rrbracket}$. \\
				\quad $\delta_B \leftarrow B - \sum_{i=1}^{N} [R_B^{[e]}]^{\llbracket i \rrbracket}$. \\
				\quad $\vec{\delta}_x \leftarrow A - \sum_{i=1}^{N} [\vec{R}_x^{[e]}]^{\llbracket i \rrbracket}$. \\
				\quad $\vec{\delta}_y \leftarrow A - \sum_{i=1}^{N} [\vec{R}_y^{[e]}]^{\llbracket i \rrbracket}$.

				\item Compute the commitments \\
				\quad $\texttt{com}^{[e]}_i := \text{Hash}_0(\texttt{salt}, e, i, [w^{[e]}_i], [g^{[e]}_i])$ \\
				for all $i \in [1:N]$, and compute the Merkle root \\
				\quad $\tilde{h}^{[e]} := \text{MerkleTree}(\texttt{com}^{[e]}_1, \dots, \texttt{com}^{[e]}_N)$.
			\end{enumerate}

			\vspace{0.5em}
			\textbf{Phase 2: First challenge (randomness for the MPC protocol).}
			\begin{enumerate}[label=\arabic*.]
				\item Compute $h_1 := \text{Hash}(m, \texttt{salt}, \tilde{h}^{[1]}, \dots, \tilde{h}^{[\tau]})$.
				\item Expand $h_1$ as $(e^{[e]})_{e \in [1:\tau]} \leftarrow \text{Expand}(h_1)$.
			\end{enumerate}

			\vspace{0.5em}
			\textbf{Phase 3: Simulation of the MPC protocol.} For each $e \in [1:\tau]$:
			\begin{enumerate}[label=\arabic*.]
				\item Computes, for $i \in S$, \\
				\quad $[\alpha^{[e]}_i] := \varphi_{x,e,f}([w^{[e]}_i], [g^{[e]}_i])$ \\
				and recomposes $\alpha^{[e]}$. \\

				\textit{This step is repeated as many times as in the MPC protocol (cf Protocol 11).}
			\end{enumerate}

			\vspace{0.5em}
			\textbf{Phase 4: Second challenge (parties to be opened).}
			\begin{enumerate}[label=\arabic*.]
				\item Compute $h_2 := \text{Hash}(\texttt{salt}, h_1, [\alpha^{[1]}]_{S}, \dots, [\alpha^{[\tau]}]_{S})$.
				\item Expand $h_2$ as $(r^{[e]})_{e \in [1:\tau]} \leftarrow \text{Expand}(h_2)$.
			\end{enumerate}

			\vspace{0.5em}
			\textbf{Phase 5: Building of the signature.} Output the signature $\sigma$ built as:

			\[
			\sigma := \texttt{salt} \parallel h_1 \parallel h_2 \parallel (([w^{[e]}_i], [g^{[e]}_i])_{i \in r^{[e]}}, \texttt{auth}^{[e]}, \alpha^{[e]})_{e \in [1:\tau]}
			\]

			where $\texttt{auth}^{[e]}$ is the authentication path for $\{\texttt{com}^{[e]}_i\}_{i \in r^{[e]}}$ w.r.t. Merkle root $\tilde{h}^{[e]}$ and $i^{[e]} \notin r^{[e]}$.
		\end{tcolorbox}

@@

	\begin{theorem}
		Suppose that no adversary can distinguish the PRF used by the protocol from real random bits
		$$
		| \mathcal{P}_{seed \in \{0,1\}^n}[b=0 | p \leftarrow PRF(seed)] - \mathcal{P}[b=1 | p \overset{{\scriptscriptstyle\$}}{\leftarrow} \{0,1\}^{2n}] | \leq \frac{1}{2}.
		$$
		Suppose also that the protocol uses a cryptographic hash function which is one-way and does not have collision attacks with better complexity than $\frac{1}{\lambda}$ using a $\mathcal{PPT}$, where $\lambda$ is the number security bits needed for the signature scheme.
		Finally, suppose that the SBC problem is hard, i.e. the complexity the best attack of the SBC problem has complexity $\frac{1}{\lambda}$ using a $\mathcal{PPT}$.
		Then, the SBC signature scheme is \textbf{EUF-CMA}.
	\end{theorem}
	\begin{proof}
		\textbf{Il suffit de modeler le schema avec les énoncés de \cite{feneu2023} et hop on peut simplement utiliser ses preuves qui sont assez gnérales pour tout système de signature MPCitH avec une oneway function et une PRF.}

	\end{proof}
